---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# secretbase

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/secretbase?color=17411d)](https://CRAN.R-project.org/package=secretbase)
[![R-multiverse status](https://img.shields.io/badge/dynamic/json?url=https%3A%2F%2Fcommunity.r-multiverse.org%2Fapi%2Fpackages%2Fsecretbase&query=%24.Version&label=r-multiverse)](https://community.r-multiverse.org/secretbase)
[![secretbase status badge](https://shikokuchuo.r-universe.dev/badges/secretbase?color=e4723a)](https://shikokuchuo.r-universe.dev/secretbase)
[![R-CMD-check](https://github.com/shikokuchuo/secretbase/workflows/R-CMD-check/badge.svg)](https://github.com/shikokuchuo/secretbase/actions)
[![codecov](https://codecov.io/gh/shikokuchuo/secretbase/graph/badge.svg)](https://app.codecov.io/gh/shikokuchuo/secretbase)
[![DOI](https://zenodo.org/badge/745691432.svg)](https://zenodo.org/doi/10.5281/zenodo.10553139)
<!-- badges: end -->

```
  ________  
 /\ sec   \
/  \ ret   \
\  /  base /
 \/_______/
```

Fast and memory-efficient streaming hash functions and base64 encoding and decoding.

Performs direct hashing of strings and raw vectors. Stream hashes files potentially larger than memory, as well as in-memory objects through R's serialization mechanism.

Implementations include the SHA-256, SHA-3 and 'Keccak' cryptographic hash functions, SHAKE256 extendable-output function (XOF), and 'SipHash' pseudo-random function.

The SHA-3 Secure Hash Standard was published by the National Institute of Standards and Technology (NIST) in 2015 at [doi:10.6028/NIST.FIPS.202](https://dx.doi.org/10.6028/NIST.FIPS.202). SHA-3 is based on the Keccak algorithm, designed by G. Bertoni, J. Daemen, M. Peeters and G. Van Assche.

The SHA-256 Secure Hash Standard was published by NIST in 2002 at <https://csrc.nist.gov/publications/fips/fips180-2/fips180-2.pdf>.

The SipHash family of pseudo-random functions by Jean-Philippe Aumasson and Daniel J. Bernstein was published in 2012 at <https://ia.cr/2012/351>.<sup>[1]</sup>

### Overview

```{r secretbase}
library(secretbase)
```

#### SHA-3
 
For the SHA-3 cryptographic hash algorithm, specify 'bits' as `224`, `256`, `384` or `512`:

```{r sha3}
sha3("secret base")
sha3("secret base", convert = FALSE)
sha3("秘密の基地の中", bits = 512L)
```

#### Hash strings and raw vectors

Character strings and raw vectors are hashed directly (as per the above).

#### Stream hash R objects

All other objects are stream hashed using R serialization

- memory-efficient as performed without allocation of the serialized object
- portable as always uses R serialization version 3 big-endian representation, skipping headers (which contain R version and native encoding information)

```{r streaming}
sha3(data.frame(a = 1, b = 2), bits = 224L)
sha3(NULL)
```

#### Stream hash files

Files are read and hashed incrementally, accepting files larger than memory:
```{r files}
file <- tempfile(); cat("secret base", file = file)
sha3(file = file)
```
```{r unlink, echo=FALSE}
unlink(file)
```

#### Hash to integer / SHAKE256 XOF

May be used as deterministic random seeds for R's pseudo random number generators (RNGs). <br />
Specify 'convert' as `NA` (and 'bits' as `32` for a single integer value):
```{r integer}
shake256("秘密の基地の中", bits = 32L, convert = NA)
```

For use in parallel computing, this is a valid method for reducing to a negligible probability that RNGs in each process may overlap. This may be especially suitable when first-best alternatives such as using recursive streams are too expensive or unable to preserve reproducibility. <sup>[2]</sup>

#### Keccak

```{r keccak}
keccak("secret base", bits = 384L)
```

#### SHA-256

```{r sha256}
sha256("secret base")
```

For a SHA-256 HMAC, pass a character string or raw vector to 'key':
```{r hmac}
sha256("secret base", key = "秘密の基地の中")
```

#### SipHash

SipHash-1-3 is optimized for performance. <br />
Pass a character string or raw vector to 'key' - up to 16 bytes (128 bits) of the key data is used:
```{r siphash}
siphash13("secret base", key = charToRaw("秘密の基地の中"))
```

#### Base64 Encoding / Decoding

Strings:
```{r base64str}
base64enc("secret base")
base64dec(base64enc("secret base"))
```
Raw vectors:
```{r base64raw}
base64enc(as.raw(c(1L, 2L, 4L)), convert = FALSE)
base64dec(base64enc(as.raw(c(1L, 2L, 4L))), convert = FALSE)
```
Serialized objects:
```{r base64ser}
base64enc(data.frame())
base64dec(base64enc(data.frame()), convert = NA)
```

### Installation

Install the latest release from CRAN or R-multiverse:

```{r cran, eval=FALSE}
install.packages("secretbase")
```

The current development version is available from R-universe:

```{r runiv, eval=FALSE}
install.packages("secretbase", repos = "https://shikokuchuo.r-universe.dev")
```

### Implementation Notes

The SHA-256, SHA-3, Keccak, and base64 implementations are based on those by the 'Mbed TLS' Trusted Firmware Project at <https://www.trustedfirmware.org/projects/mbed-tls>.

The SipHash implementation is based on that of Daniele Nicolodi, David Rheinsberg and Tom Gundersen at <https://github.com/c-util/c-siphash>, which is in turn based on the reference implementation by Jean-Philippe Aumasson and Daniel J. Bernstein released to the public domain at <https://github.com/veorq/SipHash>.

### References

[1] Jean-Philippe Aumasson and Daniel J. Bernstein (2012), *"SipHash: a fast short-input PRF"*, Paper 2012/351, Cryptology ePrint Archive, <https://ia.cr/2012/351>.

[2] Pierre L’Ecuyer, David Munger, Boris Oreshkin and Richard Simard (2017), *"Random numbers for parallel computers: Requirements and methods, with emphasis on GPUs"*, Mathematics and Computers in Simulation, Vol. 135, May 2017, pp. 3-17 [doi:10.1016/j.matcom.2016.05.00](https://doi.org/10.1016/j.matcom.2016.05.005).

### Links

◈ secretbase R package: <https://shikokuchuo.net/secretbase/>

Mbed TLS website: <https://www.trustedfirmware.org/projects/mbed-tls><br />
SipHash streaming implementation: <https://github.com/c-util/c-siphash><br />
SipHash reference implementation: <https://github.com/veorq/SipHash>

--

Please note that this project is released with a [Contributor Code of Conduct](https://shikokuchuo.net/secretbase/CODE_OF_CONDUCT.html). By participating in this project you agree to abide by its terms.
