---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# secretbase

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/secretbase?color=42147b)](https://CRAN.R-project.org/package=secretbase)
[![secretbase status badge](https://shikokuchuo.r-universe.dev/badges/secretbase?color=e4723a)](https://shikokuchuo.r-universe.dev/secretbase)
[![R-CMD-check](https://github.com/shikokuchuo/secretbase/workflows/R-CMD-check/badge.svg)](https://github.com/shikokuchuo/secretbase/actions)
[![codecov](https://codecov.io/gh/shikokuchuo/secretbase/graph/badge.svg)](https://app.codecov.io/gh/shikokuchuo/secretbase)
[![DOI](https://zenodo.org/badge/745691432.svg)](https://zenodo.org/doi/10.5281/zenodo.10553139)
<!-- badges: end -->

SHA-256, SHA-3 cryptographic hash and SHAKE256 extendable-output functions (XOF).

The SHA-3 Secure Hash Standard was published by the National Institute of Standards and Technology (NIST) in 2015 at [doi:10.6028/NIST.FIPS.202](https://dx.doi.org/10.6028/NIST.FIPS.202).

Fast and memory-efficient implementation using the core algorithms from 'Mbed TLS' under the Trusted Firmware Project <https://www.trustedfirmware.org/projects/mbed-tls/>.

### Installation

Install the latest release from CRAN:

```{r cran, eval=FALSE}
install.packages("secretbase")
```

or the development version from rOpenSci R-universe:

```{r runiv, eval=FALSE}
install.packages("secretbase", repos = "https://shikokuchuo.r-universe.dev")
```

### Quick Start

`secretbase` offers the functions: `sha3()` and `sha256()`.

For `sha3()`, to use:
 
 - SHA-3 cryptographic hash algorithm, specify 'bits' as `224`, `256`, `384` or `512`
 - SHAKE256 extendable-output function (XOF), specify any other bit length

```{r secretbase}
library(secretbase)

sha3("secret base")

sha3("secret base", convert = FALSE)

sha3("秘密の基地の中", bits = 512)

```

Hash arbitrary R objects:

 - uses memory-efficient 'streaming' serialization (no allocation of serialized object)
 - portable as always uses R serialization v3 XDR, skipping headers (containing R version and encoding information)

```{r streaming}
sha3(data.frame(a = 1, b = 2), bits = 160)

sha3(NULL)
```

Hash files:

 - in a streaming fashion, accepting files larger than memory

```{r files}
file <- tempfile(); cat("secret base", file = file)
sha3(file = file)
```
```{r unlink, echo=FALSE}
unlink(file)
```

Hash to integer:

 - specify 'convert' as `NA` (and 'bits' as `32` for a single integer value)
 - may be supplied as deterministic random seeds for R's pseudo random number generators (RNGs)

```{r integer}
sha3("秘密の基地の中", bits = 384, convert = NA)

sha3("秘密の基地の中", bits = 32, convert = NA)
```

For use in parallel computing, this is a valid method for reducing to a negligible probability that RNGs in each process may overlap. This may be especially suitable when first-best alternatives such as using recursive streams are too expensive or unable to preserve reproducibility. <sup>[1]</sup>

### References

[1] Pierre L’Ecuyer, David Munger, Boris Oreshkin and Richard Simard (2017), *"Random numbers for parallel computers: Requirements and methods, with emphasis on GPUs"*, Mathematics and Computers in Simulation, Vol. 135, May 2017, pp. 3-17 [doi:10.1016/j.matcom.2016.05.00](https://doi.org/10.1016/j.matcom.2016.05.005).

### Links

Links:

`secretbase` website: <https://shikokuchuo.net/secretbase/><br />
`secretbase` on CRAN: <https://cran.r-project.org/package=secretbase><br />

Mbed TLS website: <https://www.trustedfirmware.org/projects/mbed-tls/><br />

--

Please note that this project is released with a [Contributor Code of Conduct](https://shikokuchuo.net/secretbase/CODE_OF_CONDUCT.html). By participating in this project you agree to abide by its terms.
