---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# secretbase

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/secretbase?color=42147b)](https://CRAN.R-project.org/package=secretbase)
[![secretbase status badge](https://shikokuchuo.r-universe.dev/badges/secretbase?color=e4723a)](https://shikokuchuo.r-universe.dev/secretbase)
[![R-CMD-check](https://github.com/shikokuchuo/secretbase/workflows/R-CMD-check/badge.svg)](https://github.com/shikokuchuo/secretbase/actions)
[![codecov](https://codecov.io/gh/shikokuchuo/secretbase/graph/badge.svg)](https://app.codecov.io/gh/shikokuchuo/secretbase)
<!-- badges: end -->

Fast, dependency-free SHA-3 cryptographic hash and SHAKE256 extendable-output function (XOF) algorithms.

The SHA-3 Secure Hash Standard was published by NIST in 2015 at [doi:10.6028/NIST.FIPS.202](https://dx.doi.org/10.6028/NIST.FIPS.202).

Uses the implementation by the 'Mbed TLS' library from the Trusted Firmware Project <https://www.trustedfirmware.org/projects/mbed-tls/>.

### Installation

Install the latest release from CRAN:

```{r cran, eval=FALSE}
install.packages("secretbase")
```

or the development version from rOpenSci R-universe:

```{r runiv, eval=FALSE}
install.packages("secretbase", repos = "https://shikokuchuo.r-universe.dev")
```

### Quick Start

`secretbase` offers one main function `sha3()`:

To use the SHA-3 cryptographic hash algorithm, specify argument 'size' as one of '224', '256', '384' or '512'.

Specify an arbitrary output size to use the SHAKE256 algorithm as an extendable-output function (XOF).

```{r secretbase}
library(secretbase)

sha3("secret base")

sha3("secret base", convert = FALSE)

sha3("秘密の基地", size = 224)

sha3("秘密の基地", size = 512)
```

To hash to an integer value (for example to generate random seeds for R's pseudo RNGs), specify a size of '32' and pass the resulting raw vector to `read_integer()`.

```{r readinteger}
hash <- sha3("秘密の基地", size = 32, convert = FALSE)
hash

read_integer(hash)
```

### Links

Links:

`secretbase` website: <https://shikokuchuo.net/secretbase/><br />
`secretbase` on CRAN: <https://cran.r-project.org/package=secretbase><br />

Mbed TLS website: <https://www.trustedfirmware.org/projects/mbed-tls/><br />

--

Please note that this project is released with a [Contributor Code of Conduct](https://shikokuchuo.net/secretbase/CODE_OF_CONDUCT.html). By participating in this project you agree to abide by its terms.
