---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# secretbase

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/secretbase?color=42147b)](https://CRAN.R-project.org/package=secretbase)
[![secretbase status badge](https://shikokuchuo.r-universe.dev/badges/secretbase?color=e4723a)](https://shikokuchuo.r-universe.dev/secretbase)
[![R-CMD-check](https://github.com/shikokuchuo/secretbase/workflows/R-CMD-check/badge.svg)](https://github.com/shikokuchuo/secretbase/actions)
[![codecov](https://codecov.io/gh/shikokuchuo/secretbase/graph/badge.svg)](https://app.codecov.io/gh/shikokuchuo/secretbase)
[![DOI](https://zenodo.org/badge/745691432.svg)](https://zenodo.org/doi/10.5281/zenodo.10553139)
<!-- badges: end -->

Fast and memory-efficient streaming hash functions for strings, raw bytes, files and in-memory objects using R's serialization mechanism.

Implementations include the SHA-256 and SHA-3 cryptographic hash functions, SHAKE256 extendable-output function (XOF), and 'SipHash' pseudo-random function.

The SHA-3 Secure Hash Standard was published by the National Institute of Standards and Technology (NIST) in 2015 at [doi:10.6028/NIST.FIPS.202](https://dx.doi.org/10.6028/NIST.FIPS.202). The SHA-256 Secure Hash Standard was published by NIST in 2002 at <https://csrc.nist.gov/publications/fips/fips180-2/fips180-2.pdf>. The SipHash family of pseudo-random functions by Jean-Philippe Aumasson and Daniel J. Bernstein was published in 2012 at <https://ia.cr/2012/351>.<sup>[1]</sup>

The SHA-256 and SHA-3 implementations are based on those by the 'Mbed TLS' Trusted Firmware Project at <https://www.trustedfirmware.org/projects/mbed-tls/>. The SipHash-1-3 implementation is based on that of Daniele Nicolodi, David Rheinsberg and Tom Gundersen at <https://github.com/c-util/c-siphash>.

### Installation

Install the latest version from the R-releases R-universe, or CRAN:

```{r runiv, eval=FALSE}
install.packages("secretbase", repos = "https://r-releases.r-universe.dev")
```

Or the development version from the author's R-universe:

```{r github, eval=FALSE}
install.packages("secretbase", repos = "https://shikokuchuo.r-universe.dev")
```

### Quick Start

`secretbase` offers the functions: `sha3()`, `sha256()` and `siphash13()`.

##### SHA-3 and XOF usage:
 
 - For the SHA-3 cryptographic hash algorithm, specify 'bits' as `224`, `256`, `384` or `512`
 - For the SHAKE256 extendable-output function (XOF), specify any other bit length

```{r secretbase}
library(secretbase)

sha3("secret base")

sha3("secret base", convert = FALSE)

sha3("秘密の基地の中", bits = 512)

```

##### Hash arbitrary R objects:

 - uses memory-efficient 'streaming' serialization (no allocation of serialized object)
 - portable as always uses R serialization v3 XDR, skipping headers (containing R version and encoding information)

```{r streaming}
sha3(data.frame(a = 1, b = 2), bits = 160)

sha3(NULL)
```

##### Hash files:

 - in a streaming fashion, accepting files larger than memory

```{r files}
file <- tempfile(); cat("secret base", file = file)
sha3(file = file)
```
```{r unlink, echo=FALSE}
unlink(file)
```

##### Hash to integer:

 - specify 'convert' as `NA` (and 'bits' as `32` for a single integer value)
 - may be supplied as deterministic random seeds for R's pseudo random number generators (RNGs)

```{r integer}
sha3("秘密の基地の中", bits = 384, convert = NA)

sha3("秘密の基地の中", bits = 32, convert = NA)
```

For use in parallel computing, this is a valid method for reducing to a negligible probability that RNGs in each process may overlap. This may be especially suitable when first-best alternatives such as using recursive streams are too expensive or unable to preserve reproducibility. <sup>[2]</sup>

##### Using a keyed hash:

- Use `spihash13()` passing an atomic vector to 'key'.
- Up to 16 bytes (128 bits) are used i.e. the length of 1 complex number, 2 doubles, 4 integers, or 16 individual characters.

```{r siphash}
siphash13("secret base", key = "秘密の基地の中")

siphash13("secret base", key = 1.2 + 3.4i)
```

### References

[1] Jean-Philippe Aumasson and Daniel J. Bernstein (2012), *"SipHash: a fast short-input PRF"*, Paper 2012/351, Cryptology ePrint Archive, <https://ia.cr/2012/351>.

[2] Pierre L’Ecuyer, David Munger, Boris Oreshkin and Richard Simard (2017), *"Random numbers for parallel computers: Requirements and methods, with emphasis on GPUs"*, Mathematics and Computers in Simulation, Vol. 135, May 2017, pp. 3-17 [doi:10.1016/j.matcom.2016.05.00](https://doi.org/10.1016/j.matcom.2016.05.005).

### Links

Links:

◈ secretbase R package: <https://shikokuchuo.net/secretbase/>

Mbed TLS website: <https://www.trustedfirmware.org/projects/mbed-tls/><br />
SipHash streaming implementation: <https://github.com/c-util/c-siphash>

--

Please note that this project is released with a [Contributor Code of Conduct](https://shikokuchuo.net/secretbase/CODE_OF_CONDUCT.html). By participating in this project you agree to abide by its terms.
